---
title: "Exercise Sheet 2"
fontsize: 11pt
header-includes: \usepackage[german]{babel}
output:
  pdf_document:
    highlight: tango
  html_document: default
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # -> Should code chunks be displayed in the rendered document?
                      eval = TRUE, # -> Should R Code Chunks be executed?
                      warning = FALSE, # -> Warnings should only be set to FALSE at the end of rendering.
                      message = FALSE) # -> Notes should only be set to FALSE at the end of rendering.
```

The _European Soccer Database_ contains data on more than 25.000 national football matches from the best European leagues. The aim of this exercise is to present interesting relationships in R using explorative data analysis and visualization.

First you need to access some tables in the database. Note: You can use the `RSQLite::dbConnect()` function to do this. To access a particular database table and convert it to a `data.frame`, you can use the `tbl_df(dbGetQuery(connection, 'SELECT * FROM table_xyz'))` command.

```{r}
library(RSQLite)
library(stringr)
library(dplyr)
library(sqldf)
library(DBI)
library(ggplot2)
# connect to database
#con <- dbConnect(SQLite(), dbname = str_c(dirname(getwd()), 
#                                          "/datasets/EuropeanSoccer.sqlite"))
conn <- DBI::dbConnect(RSQLite::SQLite(),"E:/OVGU/Subject/Sem 3/Visual Analytics/VisualAnalyticsOVGU/Datasets/EuropeanSoccer.sqlite")

table <- dbListTables(conn)

# tabele queries
match <- tbl_df(dbGetQuery(conn,"SELECT * FROM Match"))
country <- tbl_df(dbGetQuery(conn,"SELECT * FROM Country"))
league <- tbl_df(dbGetQuery(conn,"SELECT * FROM League"))
player <- tbl_df(dbGetQuery(conn,"SELECT * FROM Player"))
player_attributes <- tbl_df(dbGetQuery(conn,"SELECT * FROM Player_Attributes"))
sqlite_sequence <- tbl_df(dbGetQuery(conn,"SELECT * FROM sqlite_sequence"))
team <- tbl_df(dbGetQuery(conn,"SELECT * FROM Team"))
team_attributes <- tbl_df(dbGetQuery(conn,"SELECT * FROM Team_Attributes"))

# ...
# ...
```

1. The first leagues of Spain, England, Germany and Italy are considered the four most attractive football leagues in Europe.  
    a) In which of the four leagues do on average score the most or the fewest goals per game? 
    b) Compare the average, median, standard deviation, variance, range and interquartile distance of goals scored per match between the four most attractive European leagues and the remaining leagues.

```{r}
# Solution for task 1...

# filter - https://cran.r-project.org/web/packages/crunch/vignettes/filters.html 
# filter - https://www.guru99.com/r-select-filter-arrange.html 


# mean(match[match$totalgoals == 2,]$totalgoals) - returns vector
# mean(match[match$totalgoals == 2,]["totalgoals"]) - doesnt work as it returns dataframe (as tuple) and not vector

spain = mean(match[match$id == 21518,]$totalgoals)
england = mean(match[match$id == 1729,]$totalgoals)
germany = mean(match[match$id == 7809,]$totalgoals)
italy = mean(match[match$id == 10257,]$totalgoals)



avgscore <- c(spain, england, germany, italy)
countrynames <- c("Spain", "England", "Germany", "Italy")
barplot(avgscore, names.arg = countrynames, xlab = "Country", ylab = "Average Score per game", main = "Average Scores")
#hist(avgscore,xlab = "country",col = "green",border = "red", xlim = c(0,5), ylim = c(0,2),
 #  breaks = 5)

```
```{r}

#Part b

library(ggplot2)
match3<-match
match3$country_id <- as.character(match3$country_id)
match3<-match %>% group_by(country_id)

m<-mean(match3$home_team_goal + match3$away_team_goal)
curve(dnorm(x, mean=m), col="darkblue", lwd=2, add=TRUE)

med<-median(match3$home_team_goal + match3$away_team_goal)
curve(dnorm(x, mean=med))

std<-sd(match3$home_team_goal + match3$away_team_goal)
curve(dnorm(x, sd=std),col="darkblue", lwd=2, add=TRUE)

rang = max(match3$home_team_goal + match3$away_team_goal) - min((match3$home_team_goal + match3$away_team_goal))

vari<-var(match3$home_team_goal + match3$away_team_goal)
#curve(dnorm(x, var=vari))


plot1 <- ggplot(match3, aes(x=name, y=m)) +   xlab("Country") + ylab("Average Goals") +  geom_bar(stat = "identity")

plot2 <- ggplot(match3, aes(x=name, y=med)) +   xlab("Country") + ylab("Median") +  geom_bar(stat = "identity")

plot3 <- ggplot(match3, aes(x=name, y=std)) +   xlab("Country") + ylab("Standard Deviation") +  geom_bar(stat = "identity")

plot4 <- ggplot(match3, aes(x=name, y=vari)) +   xlab("Country") + ylab("Variance") +  geom_bar(stat = "identity")

plot5 <- ggplot(match3, aes(x=name, y=rang)) +   xlab("Country") + ylab("Range") +  geom_bar(stat = "identity")

library(ggpubr)
figure <- ggarrange(plot1, plot2, plot3,plot4,plot5,
                    ncol = 2, nrow = 3)
figure

```

2. Is there really a home advantage? Use a box plot to show the number of goals scored by home and away teams.

```{r}
# Solution for task 2...
# https://www.statmethods.net/graphs/boxplot.html
teamnames <- c("home","away")
boxplot(match$home_team_goal, match$away_team_goal, data = match, xlab = "teams", names = teamnames, ylab = "away team goal ", main = "Goals scored by teams")
```

3. _"All soccer players are fair-weather players!"_ Check the assertion with a line chart: Do on average more goals fall per game in the summer months than in the rest of the year?


```{r}
# # Solution for task 3...
# # https://www.statmethods.net/graphs/line.html
# # https://www.tutorialspoint.com/r/r_line_graphs.htm 
# # https://stackoverflow.com/questions/1699046/for-each-row-in-an-r-dataframe 
# # summer months - May (05) - Aug (08)
# 
# goals <- match$home_team_goal + match$away_team_goal
# 
# months <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
# counter <- c(0,0,0,0,0,0,0,0,0,0,0,0)
# avg <- c(0,0,0,0,0,0,0,0,0,0,0,0)
# avg_old <- c(0,0,0,0,0,0,0,0,0,0,0,0)
# 
# for(j in months){
#      if( "05" == j){  
#         counter[as.numeric(j)] = counter[as.numeric(j)] +1  
#        print( counter[as.numeric(j)])
#        print(as.numeric(goals[1]))
#        avg_old = avg
#        print(avg_old)
#       # print(as.numeric(j))
#        avgfunc(counter[as.numeric(j)],as.numeric(goals[1]),as.numeric(j),avg_old)
#    }}
# 
# counter
# 
# 
# avgfunc <- function(count,goal,monthj,avg_old){
#   avg[monthj] = (((avg_old[monthj]*(count-1))+goal)/count)
#  #print(c(avg))
#  return(c(avg))
# }
# 
# for(i in 1:nrow(match)){
#   month_str <- str_sub(match$date[i],6,7)     #extract month from string
#   for(j in months){
#    if( month_str == j){  
#      counter[as.numeric(j)] = counter[as.numeric(j)] +1              #increase the counter for jth month
#      #print (counter[j])  
#       avgfunc(counter[as.numeric(j)],as.numeric(goals[i]),as.numeric(j),avg_old)          #calc incremental average for jth month
#      }
#   }
# }
# 
# 
# plot(avg,type = "o", col = "red", xlab = "Month", ylab = "Avg Goals",
#    main = "Average Goals in months")

```
3. _"All soccer players are fair-weather players!"_ Check the assertion with a line chart: Do on average more goals fall per game in the summer months than in the rest of the year?


```{r}

#Attempt 2

# dataframe - https://www.datamentor.io/r-programming/data-frame/
# select, filter - https://www.guru99.com/r-select-filter-arrange.html
# mean - https://www.tutorialspoint.com/r/r_mean_median_mode.htm 
# filter - https://dplyr.tidyverse.org/reference/filter.html 
# filter - https://stackoverflow.com/questions/1686569/filter-data-frame-rows-by-a-logical-condition 


match<- mutate(match, month = as.numeric(str_sub(match$date,6,7)), totalgoals = (match$home_team_goal + match$away_team_goal))
result<- c()
months <- seq(1:12)
for(i in months) {
  result [i] <- mean((filter(match[120],match$month == i))$totalgoals)
}

plot(result,type = "o", col = "red", xlab = "Month", ylab = "Avg Goals",
   main = "Average Goals in months")

```

4. Display the average goals scored per game for the top 4 leagues per year from 2008 to 2016.

```{r}
# Solution for task 4...

#filter(match[120], match$country_id == as.numeric(1729))
#filter(match[120], match$country_id == as.numeric(1729))$totalgoals
#1st is in DF and 2nd in List. Mean works with list

#mean(filter(match[120], match$country_id == as.numeric(1729) & match$year == 2008)$totalgoals)

match<- mutate(match, year = as.numeric(str_sub(match$date,1,4)))
#yearrange <-seq(2008:2016) not working
yearrange <- c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016)
j = 1
#spain = england = germany = italy = 0
for(i in as.numeric(yearrange)){
  spain[j] <- mean(filter(match[120], match$country_id == as.numeric(21518) & match$year == as.numeric(i))$totalgoals)
  england[j] <- mean(filter(match[120], match$country_id == as.numeric(1729) & match$year == as.numeric(i))$totalgoals)
  germany[j] <- mean(filter(match[120], match$country_id == as.numeric(7809) & match$year == as.numeric(i))$totalgoals)
  italy[j] <- mean(filter(match[120], match$country_id == as.numeric(10257) & match$year == as.numeric(i))$totalgoals)
  j = j+1 
}
spain
england
germany
italy

#barplot(avgscore, names.arg = countrynames, xlab = "Country", ylab = "Average Score per game", main = "Average Scores")
#countrynames <- c("Spain", "England", "Germany", "Italy")

# https://stackoverflow.com/questions/36968602/multiple-barplots-in-r-side-by-side
# https://stats.stackexchange.com/questions/14118/drawing-multiple-barplots-on-a-graph-in-r 


bars <- rbind(spain, england, germany, italy)
barplot(bars, beside = T,  names.arg = c(yearrange), xlab = "Year", ylab = "Average Score per game", col = terrain.colors(4), ylim = c(0,5))
legend(40, 5, c("Spain", "England", "Germany", "Italy"), cex=0.6, 
       fill=terrain.colors(4))

```

5. Use an estimated density function curve AND a QQ-Plots to check whether the `home_team_possession` variable is (approximately) normally distributed.

```{r}
# Solution for task 5...

```

6. Use a box plot to show whether there is a correlation between ball ownership (`home_team_possession`) and the number of goals (`home_team_goals`) scored per game for home teams. Create four categories of ball ownership shares: _very low_ ($\leq 25\%$), _low_  ($25\%<x\leq50\%$), _high_ ($50\%<x\leq 75\%$) und _very high_ ($x>75\%$).

```{r}
# Solution for task 6...

library(corrplot)

homegoal<-match$home_team_goal
homeposs<-match$home_team_possession


boxplot(homegoal, homeposs,
main="Multiple boxplots for comparision of goals",
names = c("Home Team Goal", "Home Team Possession"),
xlab="Teams",
ylab="Goals",
col="orange",
border="brown"
)
#v1<-mutate(goal = match$home_team_goal, poss = match$home_team_possession)
#v1<-match
#v1
#v1[1:10] <- NULL
#v1[2:106] <- NULL
#v1[3:7] <- NULL
#getting home team goal and home team possessions only


v1<-data.frame(goal=numeric(),poss=numeric())
str(match$home_team_goal)
v1$goal <- as.numeric(match$home_team_goal)
v1$poss <- as.numeric(match$home_team_possession)
v1$date<-NULL

cor_r<-cor(v1, method = "spearman", use = "complete.obs")
corrplot(cor_r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)


```

------
Dataset:

- http://isgwww.cs.uni-magdeburg.de/cv/lehre/VisAnalytics/material/exercise/datasets/EuropeanSoccer.sqlite  
(For database schema and explanation of variables, see: https://www.kaggle.com/hugomathien/soccer)